\documentclass{article}
\usepackage{graphicx} % Required for inserting images
%No tengo el archivo header-posets, uso uno mío que debe ser parecido
%\input{header-posets}
\input{header-trabajo-eng}
\title{Problema de teoría de conjuntos}


\begin{document}
	

\section*{Introduction}

Within the real numbers, there exist multiple infinitely decreasing sequences. Consider the following question: for which functions \(f\) does there exist an initial point \(x_0\) such that the sequence defined by \(x_{n+1} = f(x_n)\) is decreasing? Clearly, for some functions (such as logarithms) this is not the case.

This motivates the following question about decreasing sequences of cardinals: in ZFC, it is possible to construct decreasing sequences of cardinals, but it is not possible to do so in such a way that
\[
X_n =_c \mathcal{P}(X_{n+1}),
\]
where $=_c$ denotes that the sets are bijective. This remains true in weaker theories, since the Axiom of Choice is not required. In this article we present a proof of this fact without using the Replacement axiom.

%Encierro WO en \text{} para que quede más junto, avisame si no te gusta así
\begin{definition}\label{aleph}
    Let $X$ be any set,
    \[
    	\text{WO}(X):=\{\langle A,R\rangle\ |A\subset X \ \text{is well-ordered by}\ R\} := \text{W}, 
    \]
    and $\aleph(X):= \text{W}/{\sim_X}$ where $\langle A,R\rangle \sim_X \langle B, R'\rangle \Longleftrightarrow \langle A, R\rangle =_o \langle B, R'\rangle$ and $=_o$ denotes initial similarity (Moschovakis, def. 7.27 \cite{Moscho}).
    Given $\aleph(X)$, and $A,B \in\aleph(X)$ we shall say $A \leq_{\aleph(X)} B$ iff for any (all) $U\in A$ and $V\in B$ it holds that $U\leq_o V$, where this last relation means $U$ is isomorphic to an initial segment of $V$.
\end{definition}

\begin{lemma}\label{aleph2}
	For all $X$, $\aleph(X)$ has the following properties:
	\begin{enumerate}[(i)]
		\item $\aleph(X)\nleq_c X$
		\item $(\aleph(X),\leq_{\aleph(X)})$ is a well order and is not equinumerous with any of its initial segments.
	\end{enumerate}
\end{lemma}

%\begin{claim}
 %   $\aleph(X)$ is a cardinal and $\aleph(X)\nleq_c X$
%\end{claim}

\begin{theorem}
	For any set $X$, $\aleph(X)<_c \mathcal{P}^4(X)$.
\end{theorem}

\begin{proof}
We will show that there's an injective function from $\aleph(X)$ into $\mathcal{P}^4(X)$. Let $f:WO(X)\rightarrow \mathcal{P}^3(X)$ defined by $f(\langle A,R\rangle) = \{\{\{a\},\{a,b\}\}| a\leq_R b\}$. Note that $f$ is injective. Now, we can define an injective function $\Tilde{f}: \mathcal{P} (WO(X)) \rightarrow \mathcal{P}^4(X)$ given by $\Tilde{f}(S) = \{f(s) | s\in S\}$. This shows that $\mathcal{P} (WO(X)) \leq_c \mathcal{P}^4(X)$. Hence, since $\aleph (X) \subset \mathcal{P} (WO(X))$, we obtained what we wanted to show.
\end{proof}

\begin{theorem}\label{partes3}
	For any set $X$, $\aleph(X)<_c \mathcal{P}^3(X)$.
\end{theorem}
\begin{proof}
%	We will show that there's an injective function from $\aleph(X)$ into $\mathcal{P}^4(X)$. Let $f:WO(X)\rightarrow \mathcal{P}^3(X)$ defined by $f(\langle A,R\rangle) = \{\{\{a\},\{a,b\}\}| a\leq_R b\}$. Note that $f$ is injective. Now, we can define an injective function $\Tilde{f}: \mathcal{P} (WO(X)) \rightarrow \mathcal{P}^4(X)$ given by $\Tilde{f}(S) = \{f(s) | s\in S\}$. This shows that $\mathcal{P} (WO(X)) \leq_c \mathcal{P}^4(X)$. Hence, since $\aleph (X) \subset \mathcal{P} (WO(X))$, we obtained what we wanted to show.
%\end{proof}

Let $f:WO(X)\rightarrow \mathcal{P}^2(X)$ defined by $f(\langle A,\mathrel{R}\rangle) = \{\{b\mid b \mathrel{R} a\} \mid a\in A\}$. $f$ is clearly injective and an argument similar as before yields what we wanted to show.
\end{proof}

\begin{theorem}[ZFC]
	There is no sequence $\{X_n\}_{n\in\omega}$ such that for all $n\in\omega$ $X_{n+1} <_c X_n$.
\end{theorem}
\begin{proof}
	Assume by contradiction that there is such a sequence, and let us take $\{\alpha_n\}_{n\in\omega}$ such that each $\alpha_n$ is an ordinal with the same cardinality as $X_n$. We then have an infinite sequence of ordinals of decreasing cardinality, which leads us to a contradiction.
\end{proof}

The previous result relies heavily on Choice, in fact, given any countable partially ordered set $\langle P,< \rangle$ it is consistent with ZF for there to be a family of sets $\mathcal{F}$ such that $\langle \mathcal{F},<_c\rangle$ is isomorphic to $\langle P,< \rangle$.
%CONSEGUIR CITA DE LO ANTERIOR
%También, ver si puede valer el teorema anterior y fallar Elección, posiblemente salga de
%estudiar los modelos en los que falla.

We shall use, following the convention from \cite[Def. I.3.1]{Kunen}, $\Bst$ to mean the axioms of Extensionality, Comprehension, Pairing, Union and Powerset. Let us use $\Bsti$ to denote $\Bst$ with the axiom of Infinity.

%Ver si se puede sacar Infinito sin problemas

%Kunen da la conjunción de Partes of Reemplazo.

\begin{theorem}[$\Bst$]
	There is no sequence $\{X_n\}_{n\in\omega}$ such that for all $n\in\omega$ $\mathcal{P}(X_{n+1}) <_c X_n$.
\end{theorem}

Before moving on to the proof, let us note that Definition \ref{aleph} does not rely on any axiom outside of $\Bst$, with the caveat that without Replacement, $\aleph(X)$ cannot be proven to be isomorphic to an ordinal.

\begin{proof}
	Assume there is such a sequence, $\mathcal{S} := \{X_n\}_{n\in\omega}$, then $\Bst$ proves that $\ran(\mathcal{S})$ exists (\cite[Def. I.6.6]{Kunen}).
	Moreover, for all $X\in\ran(\mathcal{S})$: 
	\[
		\aleph(X) \subset \mathcal{P}(\WO(X)) \subseteq \mathcal{P}(\WO(\union \ran(\mathcal{S}))),
	\]
	where both inclusions follow from Definition \ref{aleph}.
	
	\noindent Thus, for all $X\in\ran(\mathcal{S})$, we have $\aleph(X)\in\mathcal{P}(\mathcal{P}(\WO(\union \ran(\mathcal{S}))))$. Hence, $\Bst$ proves the existence of the sequence $\{\aleph(X_n)\}_{n\in\omega}$, which would of course be trivial using Replacement.
	
	It is straightforward to prove that for each $n\in\omega$, $\mathcal{P}^{3}(X_{n+3}) <_c X_n$, indeed, applying the hypothesis on $\mathcal{S}$ yields:
	\begin{align*}
		&\mathcal{P}(X_{n+3}) \hspace{4.5 pt} <_c X_{n+2}
		\\ \implies &\mathcal{P}^2(X_{n+3}) \leq_c \hspace{4.5 pt} \mathcal{P}(X_{n+2}) <_c X_{n+1}
		\\ \implies &\mathcal{P}^3(X_{n+3}) \leq_c \mathcal{P}^{2}(X_{n+2}) \leq_c \mathcal{P}(X_{n+1}) <_c X_n.
	\end{align*}
	As a consequence, $\aleph(X_{n+3}) <_c \mathcal{P}^{3}(X_{n+3}) <_c X_n$, where the first inequality comes from applying Theorem \ref{partes3}.
	
	\noindent Let us now show that $\aleph(X_{n+3}) <_c \aleph(X_n)$ for all $n\in\omega$. Well ordered sets are pairwise comparable with the relation $\leq_c$, in particular: $\aleph(X_{n+3}) <_c \aleph(X_n)$ or $\aleph(X_n) \leq_c \aleph(X_{n+3})$. The latter statement cannot hold, since it would imply $\aleph(X_n) <_c X_n$, which contradicts Lemma \ref{aleph2}.
	
	Finally, $\{\aleph(X_{3n})\}_{n\in\omega}$ is a set of well ordered sets that is not well founded with the $<_c$ relation, which cannot happen under $\Bst$.
\end{proof}

The following theorem is equivalent to the previous one when using Replacement but strictly stronger when working within $\Bsti$.

\begin{theorem}[$\Bsti$]
	There cannot be a formula with one free variable 
\end{theorem}


%Voy agregando bibliografía, con el Kunen y el Jech por si llegan a hacer falta
\bibliographystyle{plain}
\bibliography{bibliografia-hartogs}
\printindex

\end{document}
